Sketch Recognition and Understanding

### 2009-09-26 trunk/draft/System.tex
after spell check

After the best fit segmentation is chosen the parameter of each primitive is computed. For example, if the stroke is represented as an ellipse, the center and axii of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and parameter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

\subsection{Structural and geometrical Features(FS1)}
 Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these 13 feature are computed.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.
		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These 12 features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}.  For $n=10$ there are 32 moment features. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.

@@

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm6.jpg.eps}
	\caption{non Lineary problem }
	\label{fig:svm6}
\end{figure}

@@

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernels are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

@@

SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset. They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. There are many other successful application SVM. They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.

@@

Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
 \\ \hline
 Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1   & Global best fitness is 0.875 and error value is 729.248312158671 
 \includegraphics[scale=0.7]{afterDefense/Stroke2Seg1labeled.jpg.eps}
 \\ \hline 
  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2   & Error value is 8.61499750333895
 \includegraphics[scale=0.7]{afterDefense/Stroke2Segmentaion.jpg.eps}
 \\ \hline 
Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2

@@

 FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
 The final Feature vector of the FS1==>
ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468

@@

FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &

@@

  \\ \hline
FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
 The convex hull    [ ]
 Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265

>>>>

After the best fit segmentation is chosen the parameter of each primitive is computed. For example, if the stroke is represented as an ellipse, the center and axis of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and parameter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

\subsection{Structural and geometrical Features(FS1)}
 Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.
		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These 12 features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}.  For $n=10$ there are 32 moment features. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.

@@

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm6.jpg.eps}
	\caption{non Linearly problem }
	\label{fig:svm6}
\end{figure}

@@

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as Hilbert Schmidt kernels, first applied to SVM in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernels are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

@@

SVM have been applied to the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset. They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. There are many other successful application SVM. They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.

@@

Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
 \\ \hline
 Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1   & Global best fitness is 0.875 and error value is 729.248312158671 
 \includegraphics[scale=0.7]{afterDefense/Stroke2Seg1labeled.jpg.eps}
 \\ \hline 
  Segmentation & PSO Segmentation AlgS2 & $P_{dp}$  & AlgS2   & Error value is 8.61499750333895
 \includegraphics[scale=0.7]{afterDefense/Stroke2Segmentaion.jpg.eps}
 \\ \hline 
Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  AlgS2 161.60692855994625    Best = AlgS2

@@

 FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
 The final Feature vector of the FS1==>
ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature Perpendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468

@@

FS3 & Moments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &

@@

  \\ \hline
FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
 The convex hull    [ ]
 Area of convex hull =  -34160.5   perimeter of hull 656.1576413855265
